# Visual Acuity 

Suppose we are interested in the hypothesis that children playing outside leads them to have better eyesight.

Consider the following population of ten children whose visual acuity we can measure. 

- Visual acuity is the decimal version of the fraction given as output in standard eye exams. 
- Someone with 20/20 vision has acuity 1.0, while someone with 20/40 vision has acuity 0.5. 
- Numbers greater than 1.0 are possible for people with better than “normal” visual acuity.

```{r make visual data}
d <- data.table(
  child = 1:10, 
  y_0 = c(1.2, 0.1, 0.5, 0.8, 1.5, 2.0, 1.3, 0.7, 1.1, 1.4), 
  y_1 = c(1.2, 0.7, 0.5, 0.8, 0.6, 2.0, 1.3, 0.7, 1.1, 1.4)
)
```

In this table: 

- `y_1` means means the measured *visual acuity* if the child were to play outside at least 10 hours per week from ages 3 to 6'  
- `y_0` means the measured *visual acuity* if the child were to play outside fewer than 10 hours per week from age 3 to age 6; 
- Both of these potential outcomes *at the child level* would be measured at the same time, when the child is 6. 

1. Compute the individual treatment effect for each of the ten children.

```{r}
d[, .(individual_treatment_effect = y_1-y_0)]
```

2. Tell a "story" that could explain this distribution of treatment effects. In particular, discuss what might cause some children to have different treatment effects than others.

Perhaps, a minority of the population have genetic characteristics that make playing outside or not playing outside beneficial. For the rest of the population playing outside or not playing outside has no effect on vision. 

3. For this population, what is the true average treatment effect (ATE) of playing outside.

```{r}
d[, .(ate = mean(y_1-y_0))]
```
**-0.03**

4. Suppose we are able to do an experiment in which we can control the amount of time that these children play outside for three years. We happen to randomly assign the odd-numbered children to treatment and the even-numbered children to control. What is the estimate of the ATE you would reach under this assignment? (Please describe your work.)

Do to the guarantees of random assignment, we can express express the aveage treatment effect as $ATE = E[Y_i(1)|D_i =1] - E[Y_i(0)|D_i = 0]$. This formula holds because under random assignment the treatment and control groups have the same expected potential outcome. 
Using the sampling framework established in the questiion, the estimate for the ATE would be:
```{r}
#select all the odd numbered children
odds = d[child %% 2 !=0] 
#select all the even numbered children
evens = d[child %%2 == 0]
#subtract their means, this is directly pulled from the formula above
odds[, .(mean(y_1))] - evens[, .(mean(y_0))]
```
To reach this estimate of ATE, I first selected all the odd numbered children and then selected all the even numbered children. I then took the mean eyesight score for both groups and subtracted the control group from the treatment group to get the ATE.

5. How different is the estimate from the truth? Intuitively, why is there a difference?

The estimated ATE is off from the actual ATE by **0.3**. Although the estimated ATE is an unbiased estimator for the actual ATE, the results in any given experiment can vary from the true ATE. In this case, our sample size is very small making errors much more likely. With a much larger sample size, the estimate would likely approach the true value.

6. We just considered one way (odd-even) an experiment might split the children. How many different ways (every possible ways) are there to split the children into a treatment versus a control group (assuming at least one person is always in the treatment group and at least one person is always in the control group)?

We can use the formula $\frac{N!}{v!(N-v)!}$ and sum the total combinations. 
```{r}
total = 0
for (i in c(1:9)) {
  total = total + (factorial(10)/(factorial(i)*factorial(10-i)))
}
total
```
**1022**

7. Suppose that we decide it is too hard to control the behavior of the children, so we do an observational study instead. Children 1-5 choose to play an average of more than 10 hours per week from age 3 to age 6, while Children 6-10 play less than 10 hours per week. Compute the difference in means from the resulting observational data.

```{r}
#first half of kids
first_half = d[child <= 5]
#second half of kids
second_half = d[child > 5]
first_half[, .(mean(y_1))] - second_half[, .(mean(y_0))]

```
**-0.54**

8. Compare your answer in (7) to the true ATE. Intuitively, what causes the difference?

There is a difference of **0.48** between the true ATE and the ATE atained with observational data. The main issue with observational data is that it allows for a plethora of other factors that influence eye sight to enter the calculation, meaning we cannot assure that the ATE is truly an unbiased estimator. In this case, the children numbered 6-10 had unusually good vision leading to the result of $-0.54$. One possible explanation is that some of these children would have had the same vision if they were in the opposite group, leading to this ATE. 

